A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-20 text/formal-defnition.tex
Revising Section 3.
Wrote a pagraph of typing rules for a multi-stage calculus.

% \AI{``Transition'' is still used.}

@@

% \TConv
As in standard dependent type systems, \TConv{} allows us to replace
the type of a term with an equivalent one.
In a type system with dependent types, this kind of rule is essential
because \AI{?} two types which have different shapes may be equivalent.
For example, when we use a vector type with its size ($\textrm{Vector}\ n$),
$\textrm{Vector}\ 5$ is equivalent to $\textrm{Vector}\ (4+1)$ obviously although they are not syntactically equivalent.

% Typing rules for a multi-stage calculus
% \AI{This paragraph is not very informative.  We shouldn't just refer readers to previous work.}
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp{} are rules for a multi-stage calculus.
The rule \TTB{}, which corresponds to brackets \AI{Sounds like the rule corresponds to brackets.  What corresponds to brackets is $\TB$}, means that if term $M$ is of type $\tau$ at stage $A\alpha$,
$\TB_\alpha M$ is of type $\TW_\alpha \tau$ at stage $A$.
The rule \TTBL{}, which corresponds to escape, is the converse of \TTB.
The rule \TGen{} for stage abstraction is straightforward.
The condition $\alpha\notin\rm{FTV}(\G)\cup\rm{FTV}(A)$ ensures that
the scope of $\alpha$ is in $M$.
The rule \TIns{} is for applications of stages to stage abstractions.
\AI{Any comments?}
The rule \TCsp is the rule for CSP, which means that
if term $M$ is of type $\tau$ at stage $A$, $\%_\alpha M$ is of type $\tau$ at stage $A\alpha$.  \AI{This does not explain what this rule means.
  It just gives how the rule is read.}

@@

@@

\begin{definition}[Evaluation Context]
  Let $B$ be either \(\varepsilon\) or a transition variable \(\beta\).
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$, is defined by the following grammar (in which $A'$ stands for a non-empty stage).
  %  $A$ is assumed to be nonempty (but $B,A'$ can be empty).
  \begin{align*}
    E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon)
    \mid E^\varepsilon_B\ M \mid v^\varepsilon\ E^\varepsilon_B \mid \TB_\alpha E^\alpha_B
    \mid \Lambda\alpha.E^\varepsilon_B \mid E^\varepsilon_B\ A                                                                                    \\
    E^{A'}_B \in ECtx^{A'}_B               & ::= \square\ (\text{if } A' = B) \mid \lambda x:\tau.E^{A'}_B \mid E^{A'}_B\ M \mid v^{A'}\ E^{A'}_B \\
                                           & \mid \TB_\alpha E^{A'\alpha}_B \mid \TBL_\alpha E^{A}_B \ (\text{where } A\alpha = A')               \\
                                           & \mid \Lambda\alpha.E^{A'}_B \mid E^{A'}_B\ A \mid \%_\alpha\ E^{A'}_B \ (\text{where } A\alpha = A')
  \end{align*}
\end{definition}
\AI{some more explanation of the syntax of evaluation contexts.}

>>>>

@@

% \TConv
As in standard dependent type systems, \TConv{} allows us to replace the type of a term with an equivalent one.
For example, when we use a vector type with its size ($\textrm{Vector}\ n$),
a value of type $\textrm{Vector}\ (4+1)$ can also has type $\textrm{Vector}\ 5$ because of \TConv{}.

% Typing rules for a multi-stage calculus
The rules \TTB, \TTBL, \TGen, \TIns, and \TCsp{} are rules for a multi-stage calculus.
\TTB and \TTBL are the same with \LTP as we explained in Section \ref{sec:informal-overview}.
The rule \TGen{} for stage abstraction is straightforward.
The condition $\alpha\notin\rm{FTV}(\G)\cup\rm{FTV}(A)$ ensures that the scope of $\alpha$ is in $M$.
The rule \TIns{} is for applications of stages to stage abstractions.
\TIns{} is mainly used when we run a code value or rename a stage variable in the body of an abstraction.
The rule \TCsp{} is the rule for CSP, which means that
if term $M$ is of type $\tau$ at stage $A$, $\%_\alpha M$ is of type $\tau$ at stage $A\alpha$.
This rule enable to embed computed values into a fragment of code.

@@

\red{ここがやばい}

@@

\begin{definition}[Evaluation Context]
  Let $B$ be either \(\varepsilon\) or a stage variable \(\beta\).
  The family of sets $ECtx^A_B$ of evaluation contexts, ranged over by $E^A_B$, is defined by the following grammar (in which $A'$ stands for a non-empty stage).
  %  $A$ is assumed to be nonempty (but $B,A'$ can be empty).
  \begin{align*}
    E^\varepsilon_B \in ECtx^\varepsilon_B & ::= \square\ (\text{if\ } B = \varepsilon)
    \mid E^\varepsilon_B\ M \mid v^\varepsilon\ E^\varepsilon_B \mid \TB_\alpha E^\alpha_B
    \mid \Lambda\alpha.E^\varepsilon_B \mid E^\varepsilon_B\ A                                                                                    \\
    E^{A'}_B \in ECtx^{A'}_B               & ::= \square\ (\text{if } A' = B) \mid \lambda x:\tau.E^{A'}_B \mid E^{A'}_B\ M \mid v^{A'}\ E^{A'}_B \\
                                           & \mid \TB_\alpha E^{A'\alpha}_B \mid \TBL_\alpha E^{A}_B \ (\text{where } A\alpha = A')               \\
                                           & \mid \Lambda\alpha.E^{A'}_B \mid E^{A'}_B\ A \mid \%_\alpha\ E^{A'}_B \ (\text{where } A\alpha = A')
  \end{align*}
\end{definition}
